Manuscript_Github_Pchabaudi_pir_TPMHeatmaps

### 2021-04-22 Manuscript_pirAnalysis_Github/Pchabaudi24h/Manuscript_Github_Pchabaudi_pir_TPMHeatmaps.Rmd
Save heatmaps and use labels

#Add chapl_aapl numbers for easier heatmap drawing later.
cir_info <- cir_info %>% 
  group_by(chapl_aapl_name) %>% 
  mutate(chapl_aapl_name_num = n_distinct(Geneid)) %>% 
  mutate(chapl_aapl_name_num = paste0(
                       chapl_aapl_name, 
                       ' (n = ',
                       chapl_aapl_name_num,
                       ')'
                     ))
```

@@

combined_data_pir <- twentyfour_avgtpm$RMT$tpm %>% 
  filter(Geneid %in% cir_housekeeping_id) %>% 
  column_to_rownames(var = 'Geneid') %>% 
  filter_all(
      any_vars(. > 1)
  )

@@

ht <- Heatmap(
  log2(1+combined_data_pir_ht[,!colnames(combined_data_pir_ht) %in% 'subfam']),
  cluster_rows = FALSE,
  cluster_columns = FALSE,
  show_row_names = FALSE,
  name = 'log(TPM+1)', 
  col = colorRamp2(c(0,
                     maxcol/4,
                     maxcol/2,
                     3*maxcol/4,
                     maxcol),
                   viridis(5)),
  na_col = 'white', 
  border = TRUE,
  border_gp = gpar(lwd = 0.1),
  column_names_gp = gpar(fontsize = 10),
  row_dend_width = unit(3, "cm"),
  width = unit(4.5, 'cm'),
  height = unit(7,'cm'),
  row_title = NULL,
  left_annotation = clade_ha,
  right_annotation = anc_ha,
  row_split = forcats::fct_shift(
    relevel(
      as.factor(combined_data_pir_ht$subfam),
      ref = 'ref')
  ),
  heatmap_legend_param = list(
    labels_gp = gpar(fontsize = 10),
    title_gp = gpar(fontsize = 10,
                    fontface = 'bold'),
    at = c(0,
           maxcol/4,
           maxcol/2,
           3*maxcol/4,
           maxcol)
  )
)

@@

draw(ht, 
     row_title_side = 'left',
     column_title_side = 'bottom',
     heatmap_legend_side = 'right',
     annotation_legend_side = 'right',
     annotation_legend_list = list(clade_lgd),
     merge_legend = TRUE)
```

@@

chapl_aapl_name_ha <- rowAnnotation(
  'sub-family' = anno_simple(
    chapl_aapl_name_info,
    col = annotation_col,
  ),
  annotation_name_gp = gpar(fontsize = 10)
)

@@

ht <- Heatmap(
  log2(1+combined_data_pir_ht[,!colnames(combined_data_pir_ht) %in% 'chapl_aapl_name']),
  cluster_rows = FALSE,
  cluster_columns = FALSE,
  show_row_names = FALSE,
  name = 'log(TPM+1)', 
  col = colorRamp2(c(0,
                     maxcol/4,
                     maxcol/2,
                     3*maxcol/4,
                     maxcol),
                   viridis(5)),
  na_col = 'white', 
  border = TRUE,
  border_gp = gpar(lwd = 0.1),
  column_names_gp = gpar(fontsize = 10),
  row_dend_width = unit(3, "cm"),
  width = unit(4.5, 'cm'),
  height = unit(7,'cm'),
  row_title = NULL,
  left_annotation = chapl_aapl_name_ha,
  row_split =
    factor(combined_data_pir_ht$chapl_aapl_name,
              levels = ordering(combined_data_pir_ht$chapl_aapl_name,
                                order_vector = 'ChAPL')),
  heatmap_legend_param = list(
    labels_gp = gpar(fontsize = 10),
    title_gp = gpar(fontsize = 10,
                    fontface = 'bold'),
    at = c(0,
           maxcol/4,
           maxcol/2,
           3*maxcol/4,
           maxcol)
  ),
  row_gap = unit(c(rep.int(1, times = 2),
                   3,
                   rep.int(1, times = 4)), "mm")
)

chapl_aapl_name_lgd <- Legend(
  title = 'sub-family',
  at = str_subset(
    sort(
      factor(labels_chapl_aapl_name,
             ordering(labels_chapl_aapl_name,
                      order_vector = 'ChAPL'))
    ),
    pattern = 'NA',
    negate = TRUE
  ),
  legend_gp = gpar(fill = na.omit(annotation_col))
)

draw(ht, 
     row_title_side = 'left',
     column_title_side = 'bottom',
     heatmap_legend_side = 'right',
     annotation_legend_side = 'right',
     annotation_legend_list = list(chapl_aapl_name_lgd),
     merge_legend = TRUE)
```

```{r}
ht <- Heatmap(
  log2(1+combined_data_pir_ht[,!colnames(combined_data_pir_ht) %in% 'chapl_aapl_name']),
  cluster_rows = TRUE,
  cluster_columns = FALSE,
  show_row_names = FALSE,
  name = 'log(TPM+1)', 
  col = colorRamp2(c(0,
                     maxcol/4,
                     maxcol/2,
                     3*maxcol/4,
                     maxcol),
                   viridis(5)),
  na_col = 'white', 
  border = TRUE,
  border_gp = gpar(lwd = 0.1),
  column_names_gp = gpar(fontsize = 10),
  row_dend_width = unit(3, "cm"),
  width = unit(4.5, 'cm'),
  height = unit(7,'cm'),
  row_title = NULL,
  left_annotation = chapl_aapl_name_ha,
  # row_split = 
  #   factor(combined_data_pir_ht$chapl_aapl_name,
  #             levels = ordering(combined_data_pir_ht$chapl_aapl_name,
  #                               order_vector = 'ChAPL')),
  row_km = 9,
  heatmap_legend_param = list(
    labels_gp = gpar(fontsize = 10),
    title_gp = gpar(fontsize = 10,
                    fontface = 'bold'),
    at = c(0,
           maxcol/4,
           maxcol/2,
           3*maxcol/4,
           maxcol)
  )
)

chapl_aapl_name_lgd <- Legend(
  title = 'sub-family',
  at = str_subset(
    sort(
      factor(labels_chapl_aapl_name,
             ordering(labels_chapl_aapl_name,
                      order_vector = 'ChAPL'))
    ),
    pattern = 'NA',
    negate = TRUE
  ),
  legend_gp = gpar(fill = na.omit(annotation_col))
)

draw(ht, 
     row_title_side = 'left',
     column_title_side = 'bottom',
     heatmap_legend_side = 'right',
     annotation_legend_side = 'right',
     annotation_legend_list = list(chapl_aapl_name_lgd),
     merge_legend = TRUE)
```

>>>>

#Add chapl_aapl/subfam numbers for easier heatmap drawing later.
cir_info <- cir_info %>% 
  group_by(chapl_aapl_name) %>% 
  mutate(chapl_aapl_name_num = n_distinct(Geneid)) %>% 
  mutate(
    chapl_aapl_name_num = paste0(
      chapl_aapl_name, 
      ' (n = ',
      chapl_aapl_name_num,
      ')'
    )) %>% 
  group_by(subfam) %>% 
  mutate(subfam_num = n_distinct(Geneid)) %>% 
  mutate(
    subfam_num = paste0(
      subfam, 
      ' (n = ',
      subfam_num,
      ')'
    )
  )
```

@@

combined_data_pir <- twentyfour_avgtpm$RMT$tpm %>% 
  filter(Geneid %in% cir_housekeeping_id) %>% 
  column_to_rownames(var = 'Geneid') %>% 
  filter_all(
      any_vars(. > 10)
  )

@@

ht <- Heatmap(
  log2(1+combined_data_pir_ht[,!colnames(combined_data_pir_ht) %in% 'subfam']),
  cluster_rows = FALSE,
  cluster_columns = FALSE,
  show_row_names = FALSE,
  name = 'log(TPM+1)', 
  col = colorRamp2(c(0,
                     maxcol/4,
                     maxcol/2,
                     3*maxcol/4,
                     maxcol),
                   viridis(5)),
  na_col = 'white', 
  border = TRUE,
  border_gp = gpar(lwd = 0.1),
  column_names_gp = gpar(fontsize = 10),
  row_dend_width = unit(3, "cm"),
  width = unit(4.5, 'cm'),
  height = unit(7,'cm'),
  # row_title = NULL,
  # left_annotation = clade_ha,
  # right_annotation = anc_ha,
  row_title_rot = 0,
  row_title_gp = gpar(fontsize = 10),
  row_split = forcats::fct_shift(
    relevel(
      as.factor(combined_data_pir_ht$subfam),
      ref = 'ref')
  ),
  heatmap_legend_param = list(
    labels_gp = gpar(fontsize = 10),
    title_gp = gpar(fontsize = 10,
                    fontface = 'bold'),
    at = c(0,
           maxcol/4,
           maxcol/2,
           3*maxcol/4,
           maxcol)
  )
)

@@

draw(ht, 
     row_title_side = 'left',
     column_title_side = 'bottom',
     # heatmap_legend_side = 'right',
     # annotation_legend_side = 'right',
     # annotation_legend_list = list(clade_lgd),
     merge_legend = TRUE)

if(global_save_var){
  pdf(file = 'Manuscript_Fig_subfam_heatmap.pdf',
      height = 5,
      width = 5)
draw(ht, 
     row_title_side = 'left',
     column_title_side = 'bottom',
     heatmap_legend_side = 'right',
     # annotation_legend_side = 'right',
     # annotation_legend_list = list(clade_lgd),
     merge_legend = TRUE)
  dev.off()
}
```

@@

# chapl_aapl_name_ha <- rowAnnotation(
#   'CHAPL/AAPL' = anno_simple(
#     chapl_aapl_name_info,
#     col = annotation_col,
#   ),
#   annotation_name_gp = gpar(fontsize = 10)
# )

@@

ht <- Heatmap(
  log2(1+combined_data_pir_ht[,!colnames(combined_data_pir_ht) %in% 'chapl_aapl_name']),
  cluster_rows = FALSE,
  cluster_columns = FALSE,
  show_row_names = FALSE,
  name = 'log(TPM+1)', 
  col = colorRamp2(c(0,
                     maxcol/4,
                     maxcol/2,
                     3*maxcol/4,
                     maxcol),
                   viridis(5)),
  na_col = 'white', 
  border = TRUE,
  border_gp = gpar(lwd = 0.1),
  column_names_gp = gpar(fontsize = 10),
  row_dend_width = unit(3, "cm"),
  width = unit(4.5, 'cm'),
  height = unit(7,'cm'),
  # row_title = NULL,
  row_title_rot = 0,
  row_title_gp = gpar(fontsize = 10),
  row_split = 
    factor(combined_data_pir_ht$chapl_aapl_name,
              levels = ordering(combined_data_pir_ht$chapl_aapl_name,
                                order_vector = 'ChAPL')),
  heatmap_legend_param = list(
    labels_gp = gpar(fontsize = 10),
    title_gp = gpar(fontsize = 10,
                    fontface = 'bold'),
    at = c(0,
           maxcol/4,
           maxcol/2,
           3*maxcol/4,
           maxcol)
  ),
  row_gap = unit(c(rep.int(1, times = 2),
                   3,
                   rep.int(1, times = 4)), "mm")
)

# chapl_aapl_name_lgd <- Legend(
#   title = 'CHAPL/AAPL',
#   at = str_subset(
#     sort(
#       factor(labels_chapl_aapl_name,
#              ordering(labels_chapl_aapl_name,
#                       order_vector = 'ChAPL'))
#     ),
#     pattern = 'NA',
#     negate = TRUE
#   ),
#   legend_gp = gpar(fill = na.omit(annotation_col))
# )

if(global_save_var){
  pdf(file = 'Manuscript_SuppFig_ChAPL-AAPL_heatmap.pdf',
      height = 5,
      width = 5)
  draw(ht, 
       row_title_side = 'left',
       column_title_side = 'bottom',
       # heatmap_legend_side = 'right',
       # annotation_legend_side = 'right',
       # annotation_legend_list = list(chapl_aapl_name_lgd),
       # merge_legend = TRUE
  )
  dev.off()
}
```

```{r}
chapl_aapl_name_ha <- rowAnnotation(
  'CHAPL/AAPL' = anno_simple(
    chapl_aapl_name_info,
    col = annotation_col,
  ),
  annotation_name_gp = gpar(fontsize = 10)
)

chapl_aapl_name_lgd <- Legend(
  title = 'CHAPL/AAPL',
  at = str_subset(
    sort(
      factor(labels_chapl_aapl_name,
             ordering(labels_chapl_aapl_name,
                      order_vector = 'ChAPL'))
    ),
    pattern = 'NA',
    negate = TRUE
  ),
  legend_gp = gpar(fill = na.omit(annotation_col))
)

ht <- Heatmap(
  log2(1+combined_data_pir_ht[,!colnames(combined_data_pir_ht) %in% 'chapl_aapl_name']),
  cluster_rows = TRUE,
  cluster_columns = FALSE,
  show_row_names = FALSE,
  name = 'log(TPM+1)', 
  col = colorRamp2(c(0,
                     maxcol/4,
                     maxcol/2,
                     3*maxcol/4,
                     maxcol),
                   viridis(5)),
  na_col = 'white', 
  border = TRUE,
  border_gp = gpar(lwd = 0.1),
  column_names_gp = gpar(fontsize = 10),
  row_dend_width = unit(3, "cm"),
  width = unit(4.5, 'cm'),
  height = unit(7,'cm'),
  row_title = NULL,
  left_annotation = chapl_aapl_name_ha,
  # row_split = 
  #   factor(combined_data_pir_ht$chapl_aapl_name,
  #             levels = ordering(combined_data_pir_ht$chapl_aapl_name,
  #                               order_vector = 'ChAPL')),
  row_km = 9,
  heatmap_legend_param = list(
    labels_gp = gpar(fontsize = 10),
    title_gp = gpar(fontsize = 10,
                    fontface = 'bold'),
    at = c(0,
           maxcol/4,
           maxcol/2,
           3*maxcol/4,
           maxcol)
  )
)

draw(ht, 
     row_title_side = 'left',
     column_title_side = 'bottom',
     heatmap_legend_side = 'right',
     annotation_legend_side = 'right',
     annotation_legend_list = list(chapl_aapl_name_lgd),
     merge_legend = TRUE)

if(global_save_var){
  pdf(file = 'Interest_chapl-aapl_clustered_heatmap.pdf',
      height = 5,
      width = 5)
draw(ht, 
     row_title_side = 'left',
     column_title_side = 'bottom',
     heatmap_legend_side = 'right',
     annotation_legend_side = 'right',
     annotation_legend_list = list(chapl_aapl_name_lgd),
     merge_legend = TRUE)
  dev.off()
}
```
